.MB +5
.MT -3
.LL 65
.PN 96
.HE MP/M User's Guide
.FT   (All Information Herein is Proprietary to Digital Research.)
.sp
.pp
3.2 Basic I/O System Entry Points
.pp
The entry points into the BIOS from the cold start loader and BDOS are
detailed below.  Entry to the BIOS is through a "jump vector" located at
the base of the BIOS, as shown below (see Appendix I as well).
The jump vector is a sequence of 17 jump
instructions which send program control to the individual BIOS
subroutines.  The BIOS subroutines may be empty for certain functions
(i.e., they may contain a single RET operation) during regeneration of
MP/M, but the entries must be present in the jump vector.
The extended I/O system entry points (XIOS) immediately follow the
last BIOS entry point.
.pp
The jump vector takes the form shown below, where the
individual jump addresses are given to the left:
.li

   BIOS+00H    JMP BOOT         ; COLD START
   BIOS+03H    JMP WBOOT        ; WARM START
   BIOS+06H    JMP CONST        ; CHECK FOR CONSOLE CHAR READY
   BIOS+09H    JMP CONIN        ; READ CONSOLE CHARACTER IN
   BIOS+0CH    JMP CONOUT       ; WRITE CONSOLE CHARACTER OUT
   BIOS+0FH    JMP LIST         ; WRITE LISTING CHARACTER OUT
   BIOS+12H    JMP PUNCH        ; WRITE CHARACTER TO PUNCH DEVICE
   BIOS+15H    JMP READER       ; READ READER DEVICE
   BIOS+18H    JMP HOME         ; MOVE TO TRACK 00
   BIOS+1BH    JMP SELDSK       ; SELECT DISK DRIVE
   BIOS+1EH    JMP SETTRK       ; SET TRACK NUMBER
   BIOS+21H    JMP SETSEC       ; SET SECTOR NUMBER
   BIOS+24H    JMP SETDMA       ; SET DMA ADDRESS
   BIOS+27H    JMP READ         ; READ SELECTED SECTOR
   BIOS+2AH    JMP WRITE        ; WRITE SELECTED SECTOR
   BIOS+2DH    JMP LISTST       ; RETURN LIST STATUS
   BIOS+30H    JMP SECTRAN      ; SECTOR TRANSLATE SUBROUTINE

.pp
Each jump address corresponds to a particular subroutine which performs
the specific function, as outlined below.  There are three major
divisions in the jump table:  the system (re)initialization which
results from calls on BOOT and WBOOT, simple character I/O performed by
calls on CONST, CONIN, CONOUT, LIST, and LISTST, and diskette
I/O performed by calls on HOME, SELDSK, SETTRK, SETSEC, SETDMA, READ,
WRITE, and SECTRAN.
.pp
All simple character I/O operations are assumed to be performed in
ASCII, upper and lower case, with high order (parity bit) set to zero.
An end-of-file condition for an input device is given by an ASCII
control-z (1AH).  Peripheral devices are seen by MP/M as "logical"
devices, and are assigned to physical devices within the BIOS.
.pp
In order to operate, the BDOS needs only the CONST, CONIN, and CONOUT
subroutines (LIST and LSTST may be used by PIP, but not the
BDOS).
.cp 4
.sp
The characteristics of each device are
.sp
.in 16
.ti 4
CONSOLE    The principal interactive consoles which communicate with
the operators, accessed through CONST, CONIN, and CONOUT. Typically,
CONSOLEs are devices such as CRTs or Teletypes.
.sp
.ti 4
LIST       The principal listing device, if it exists on your system,
which is usually a hard-copy device, such as a printer or Teletype.
.sp
.ti 4
DISK       Disk I/O is always performed through a sequence of calls on the various
disk access subroutines which set up the disk number to access, the
track and sector on a particular disk, and the direct memory access
(DMA) address involved in the I/O operation.  After all these
parameters have been set up, a call is made to the READ or WRITE
function to perform the actual I/O operation.  Note that there is often
a single call to SELDSK to select a disk drive, followed by a number of
read or write operations to the selected disk before selecting another
drive for subsequent operations.  Similarly, there may be a single call
to set the DMA address, followed by several calls which read or write
from the selected DMA address before the DMA address is changed.  The
track and sector subroutines are always called before the READ or WRITE
operations are performed.
.pp
Note that the READ and WRITE routines should perform several retries
(10 is standard) before reporting the error condition to the BDOS.  If
the error condition is returned to the BDOS, it will report the error
to the user.  The HOME subroutine may or may not actually perform the
track 00 seek, depending upon your controller characteristics; the
important point is that track 00 has been selected for the next
operation, and is often treated in exactly the same manner as SETTRK
with a parameter of 00.
.pp
.in 0
The exact responsibilities of each entry point subroutine are
given below:
.sp
.sp
.in 16
.ti 4
BOOT        The BOOT entry point gets called from the MP/M loader
after it has been loaded by the cold start
loader and is responsible for basic system initialization.
Note that under MP/M a return must be made from BOOT to continue
execution of the MP/M loader.
.sp
.ti 4
WBOOT       The WBOOT entry point performs a BDOS system reset,
terminating the calling process.
.sp
.ti 4
CONST       Sample the status of the console device specified by
register D
and return 0FFH in register A if a character is ready to read,
or 00H in register A if no console characters are ready.
.sp
.ti 4
CONIN       Read the next character from the console device specified
by register D into register A, and set
the parity bit (high order bit) to zero.  If no console character is
ready, wait until a character is typed before returning.
.sp
.ti 4
CONOUT      Send the character from register C to the console output
device specified by register D.  The character is in ASCII, with high
order parity bit set to
zero.  You may want to include a delay on a line feed or carriage
return, if your console device requires some time interval at the end
of the line (such as a TI Silent 700 terminal).  You can, if you
wish, filter out control characters which cause your console device
to react in a strange way (a control-z causes the Lear Seigler
terminal to clear the screen, for example).
.sp
.ti 4
LIST        Send the character from register C to the
listing device.  The character is in ASCII with zero parity.
.sp
.ti 4
PUNCH       The punch device is not implemented under MP/M.  The
transfer vector position is preserved to maintain CP/M compatibility.
Note that MP/M supports up to 16 character I/O devices, any of which
can be a reader/punch.
.sp
.ti 4
READER      The reader device is not implemented under MP/M.  See the
note above for PUNCH.
.sp
.ti 4
HOME        Return the disk head of the currently selected disk
(initially disk A) to the track 00 position.  If your controller
allows access to the track 0 flag from the drive, step the head until
the track 0 flag is detected.  If your controller does not support
this feature, you can translate the HOME call into a call on SETTRK
with a parameter of 0.
.sp
.ti 4
SELDSK      Select the disk drive given by register C for further
operations, where register C contains 0 for drive A, 1 for drive
B, and so-forth up to 15 for drive P (the
standard MP/M distribution version supports four
drives).
On each disk select, SELDSK must return in HL the base address of a
16-byte area, called the Disk Parameter Header, described in
the digital research document titled "CP/M 2.0 Alteration Guide".
For standard floppy disk drives, the
contents of the header and associated tables does not change,
and thus the program segment included in the sample XIOS
performs this operation automatically.  If there is an attempt
to select a non-existent drive, SELDSK returns HL=0000H as
an error indicator.
.pp
On  entry to SELDSK it is possible to determine whether  it
is  the  first time the specified disk has  been  selected.
Register E, bit 0 (least significant bit) is a zero if the
drive  has not been previously selected.  This information
is  of  interest in systems which read  configuration
information  from  the disk in order to set up  a  dynamic
disk definition table.
.pp
Although SELDSK must return the header
address on each call, it is
advisable to postpone the actual physical disk select operation until an I/O
function (seek, read or write) is actually performed, since disk
selects often occur without utimately performing any disk I/O, and
many controllers will unload the head of the current disk before
selecting the new drive.  This would cause an excessive amount of
noise and disk wear.
.sp
.ti 4
SETTRK      Register BC contains the track number for subsequent disk
accesses on the currently selected drive.  You can choose to seek the
selected track at this time, or delay the seek until the next read or
write actually occurs.  Register BC can take on values in the range
0-76 corresponding to valid track numbers
for standard floppy disk drives, and 0-65535 for non-standard
disk subsystems.
.sp
.ti 4
SETSEC      Register BC contains the sector number (1 through 26) for
subsequent disk accesses on the currently selected drive.  You can
choose to send this information to the controller at this point, or
instead delay sector selection until a read or write operation occurs.
.sp
.ti 4
SETDMA      Register BC contains the DMA (disk memory access) address for
subsequent  read or write operations.  For example, if B = 00H and C
= 80H when SETDMA is called, then all subsequent read operations read
their data into 80H through 0FFH, and all subsequent write operations
get their data from 80H through 0FFH, until the next call to SETDMA
occurs.  The initial DMA address is assumed to be 80H.  Note that the
controller need not actually support direct memory access.  If, for
example, all data is received and sent through I/O ports, the XIOS
which you construct will use the 128 byte area starting at the
selected DMA address for the memory buffer during the following read
or write operations.
.sp
.ti 4
READ        Assuming the drive has been selected, the track has been
set, the sector has been set, and the DMA address has been
specified, the READ subroutine attempts to read one sector based
upon these parameters, and returns the following error codes in
register A:
.sp
.li
                 0      no errors occurred
                 1      non-recoverable error condition occurred

.br
Currently, MP/M responds only to a zero or non-zero value as the
return code.  That is, if the value in register A is 0 then MP/M
assumes that the disk operation completed properly. If an error
occurs, however, the XIOS should attempt at least 10 retries to see
if the error is recoverable. When an error is reported the BDOS will
print the message "BDOS ERR ON x:  BAD SECTOR".  The operator then
has the option of typing <cr> to ignore the error, or ctl-C to abort.
.sp
.ti 4
WRITE       Write the data from the currently selected DMA address
to the currently selected drive, track, and sector.  The data should
be marked as "non deleted data" to maintain compatibility  with other
MP/M systems.  The error codes given in the READ command are returned
in register A, with error recovery attempts as described above.
.sp
.ti 4
LISTST      Return the ready status of the list device.
The value 00 is returned in A if the list device
is not ready to accept a character, and 0FFH if a character
can be sent to the printer.  Note that a 00 value always
suffices.
.sp
.ti 4
SECTRAN     Performs sector logical to physical sector translation
in order to improve the overall response of MP/M.  Standard MP/M
systems are shipped with a "skew factor" of 6, where six physical
sectors are skipped between each logical read operation.  This
skew factor allows enough time between sectors for most programs
to load their buffers without missing the next sector.  In particular
computer systems which use fast processors, memory, and disk
subsystems, the skew factor may be changed to improve overall
response.  Note, however, that you should maintain a single
density IBM compatible version of MP/M for information transfer into and
out of your computer system, using a skew factor of 6.
In general, SECTRAN receives a logical sector number in BC, and
a translate table address in DE.  The sector number is used as
an index into the translate table, with the resulting physical
sector number in HL.  For standard systems, the tables and
indexing code is provided in the XIOS and need not be changed.
.qi
.br
