             D   I   G   I   T   A   L      R   E   S   E   A   R   C   H
     
                P. O. Box 579, Pacific Grove, CA 93950, (408) 649-3896
     
     
                                *** MP/M 1.0 ***
     
                              Application Note #02
     
     
                      Preparing a Custom Cold Start Loader
                      ------------------------------------
     
     
           The MP/M 1.0 cold start loader is usually placed on the first
       two tracks of a diskette.  It contains a boot program followed by
       the MP/M loader.  The MP/M loader is the same exact code as the
       MPMLDR.COM file.  MP/M is distributed with a cold start loader that
       is configured for a four drive single density or mixed double/single
       density Intel MDS-800.
     
           The following steps can be used to prepare a custom cold start
       loader:
     
           1.)  Prepare the MPMLDR.COM file by integrating your custom
       LDRBIOS as described in Application Note #01.  Test the MPMLDR.COM
       and verify that it operates properly.
     
           2.)  Execute either DDT or SID.
     
               A>DDT
               DDT VERS 2.0
     
           3.)  Using the input command ('I') specify that the MPMLDR.COM
       file is to be read in and then read ('R') in the file with an
       offset of 880H bytes.
     
               -IMPMLDR.COM
               -R880
               NEXT  PC
               2480 0100
     
           4.)  Using the 'I' command specify that the BOOT.HEX file is to
       be read in and then read in the file with an offset that will load
       the boot into memory at 900H.  The 'H' command can be used to
       calculate the offset.
     
               -H900 3000
               3900 D900

               -IBOOT.HEX
               -RD900
               NEXT  PC
               2480 0000
     







       MP/M 1.0  Application Note #02  (Cont'd)
   
   
           5.)  Return to the CP/M console command processor (CCP) by
       jumping to location zero and then use the SYSGEN program to write
       the new cold start loader onto the first two tracks of the diskette.
     
               -G0
     
           6.)  Use the SYSGEN program to write the new cold start loader
       onto the first two tracks of the diskette.
   
               A>SYSGEN
               SYSGEN VER 2.0
               SOURCE DRIVE NAME (OR RETURN TO SKIP)<cr>
               DESTINATION DRIVE NAME (OR RETURN TO REBOOT)B
               DESTINATION ON B, THEN TYPE RETURN<cr>
               FUNCTION COMPLETE


