             D   I   G   I   T   A   L      R   E   S   E   A   R   C   H
     
                P. O. Box 579, Pacific Grove, CA 93950, (408) 649-3896
     
     
                                *** MP/M 1.0 ***
     
                              Application Note #01
     
     
                  Integrating a custom LDRBIOS into MPMLDR.COM
                  --------------------------------------------
     
     
           The MP/M 1.0 loader resides in a file named MPMLDR.COM.  This
       loader contains a BIOS called the LDRBIOS org'd at 1700H.  MP/M is
       distributed with an MPMLDR.COM that is configured for a four drive
       single density or mixed double/single Intel MDS-800.  This MPMLDR.COM
       can be configured for a different hardware environment by coding
       a LDRBIOS as described in section 3.1 of the "MP/M User's Guide".
     
           The following steps can be used to integrate a custom LDRBIOS
       into the MPMLDR.COM:
     
           1.)  Obtain access to CP/M version 1.4 or 2.0 and prepare the
       LDRBIOS.HEX file.
     
           2.)  Read the MPMLDR.COM file into memory using either DDT
       or SID.
     
               A>DDT MPMLDR.COM
               DDT VERS 2.0
               NEXT  PC
               1A00 0100
     
           3.)  Using the input command ('I') specify that the LDRBIOS.HEX
       file is to be read in and then read ('R') in the file.  The effect
       of this operation is to overlay the BIOS portion of the MP/M
       loader.
     
               -ILDRBIOS.HEX
               -R
               NEXT  PC
               1A00 0000
     
           4.)  Return to the CP/M console command processor (CCP) by
       executing a jump to location zero.
     
               -G0
     
           5.)  Write the updated memory image onto a disk file using
       the CP/M 'SAVE' command.  The 'X' placed in front of the file
       name is used simply to designate an experimental version, preserving
       the orginal.
     
               A>SAVE 26 XMPMLDR.COM
     
           6.)  Test XMPMLDR.COM and then rename it to MPMLDR.COM.
