             D   I   G   I   T   A   L      R   E   S   E   A   R   C   H
     
                P. O. Box 579, Pacific Grove, CA 93950, (408) 649-3896
     
     
                                *** MP/M 1.0 ***
     
                              Application Note #05
     
     
                 Debugging an XIOS or a Resident System Process
                 ----------------------------------------------
     
     
           An XIOS or a resident system process can be debugged using
       DDT or SID running under CP/M 1.4 or 2.0.  The debugging
       technique is outlined in the following steps:
   
           1.)  Using GENSYS running under either CP/M or MP/M generate
       a MPM.SYS file which specifies the top of memory as 'C0'H.
   
               ...
               Top page of memory = C0
               ...
   
               Also while executing GENSYS specify the breakpoint
       restart number as that used by the CP/M SID or DDT which you
       will be executing.  This restart is usually #7.
   
               ...
               Breakpoint RST #  = 7
               ...
   
           2.)  If a resident system process is being debugged make
       certain that it is selected for inclusion in MPM.SYS.
   
           3.)  Using CP/M 1.4 or 2.0, load the MPMLDR.COM file into
       memory.
   
               A>DDT MPMLDR.COM
               DDT VERS 2.0
               NEXT  PC
               1A00 0100
   
           4.)  Place a 'B' character into the second position of
       default FCB.  This operation can be done with the 'I' command:
   
               -IB
   
           5.)  Execute the MPMLDR.COM program by entering a 'G'
       command:
   
               -G
   
           6.)  At point the MP/M loader will load the MP/M operating
       system into memory, displaying a memory map.
   
   
   
   
   
   
       MP/M 1.0  Application Note #05  (Cont'd)
   
   
           7.)  If you are debugging an XIOS, note the address of the
       XIOS.SPR memory segment.  If you are debugging a resident system
       process, note the address of the resident system process.  This
       address is the relative 0000H address of the code being debugged.
       You must also note the address of SYSTEM.DAT.
   
           8.)  Using the 'S' command, set the byte at SYSTEM.DAT + 2
       to the restart number which you want the MP/M debugger to use.
       Do not select the same restart as that being used by the CP/M
       debugger.
   
               ...
               Memory Segment Table:
               SYSTEM  DAT  C000H  0100H
               ...
   
               -SC002
               C002 07 05
   
           9.)  Using the 'X' command, determine the MP/M beginning
       execution address.  The address is the first location past
       the current program counter.
   
               -X
               ....................... P = 08CD .....
   
          10.)  Begin execution of MP/M using the 'G' command, specifying
       any breakpoints which you need in your code.  Actual memory address
       can be determined using the 'H' command to add the code segment
       base address given in the memory map to the relative displacement
       address in your XIOS or resident system process listing.
   
                The following example shows how to set a breakpoint to
       debug an XIOS list subroutine:
   
               Given the following memory map-
   
               XIOS    SPR  B800H  0500H
   
               Begin MP/M execution with the following breakpoint-
   
               -G8CE,B80F
   
          11.)  At this point you have MP/M running with CP/M and its
       debugger also in memory.  Since interrupts are left enabled during
       operation of the CP/M debugger when a breakpoint is reached, care
       must be taken to ensure that interrupt driven code does not
       execute through the point at which you have broken.

