.MB +5
.MT -3
.LL 65
.BP
.OP
.NA
.SP 7
.CE
MP/M
.SP 4
.CE
Multi-Programming Monitor Control Program
.SP 4
.CE
USER'S GUIDE
.SP 9
.BP
.AD
.SP 4
.IN 6
.LL -6
.CE
COPYRIGHT
.SP 1
Copyright (c) 1979, 1980 by Digital Research.  All rights
reserved.  No part of this publication may be reproduced,
transmitted, transcribed, stored in a retrieval system, or
translated into any language or computer language, in any
form or by any means, electronic, mechanical, magnetic,
optical, chemical, manual or otherwise, without the prior
written permission of Digital Research, Post Office Box 579,
Pacific Grove, California 93950.
.SP
This manual is, however, tutorial in nature .  Thus, permission
is granted to reproduce or abstract the example programs shown
in enclosed figures for the purposes of inclusion within the
reader's programs.
.SP
.CE
DISCLAIMER
.SP
Digital Research makes no representations or warranties with
respect to the contents hereof and specifically disclaims any
implied warranties of merchantability or fitness for any
particular purpose.  Further, Digital Research reserves the
right to revise this publication and to make changes from
time to time in the content hereof without obligation of
Digital Research to notify any person of such revision or
changes.
.SP 1
.CE
TRADEMARKS
.SP 1
CP/M is a registered trademark of Digital Research.  MP/M, MAC,
and SID are trademarks of Digital Research.
.sp 4
The "MP/M User's Guide" was prepared using the Digital Research
TEX text formatter.
.sp 3
.ce
**************************************
.ce
*  Second Printing:  February, 1980  *
.ce
**************************************
.QI
.LL +6
.HE MP/M User's Guide
.BP
.CE
Table of Contents
.LI



1.   MP/M Features and Facilities . . . . . . . . . . . . . .  1

     1.1   Introduction . . . . . . . . . . . . . . . . . . .  1
     1.2   Functional Description of MP/M . . . . . . . . . .  3
     1.3   Console Commands . . . . . . . . . . . . . . . . .  4
     1.4   Commonly Used System Programs  . . . . . . . . . .  8
     1.5   Standard Resident System Processes . . . . . . . . 13

2.   MP/M Interface Guide . . . . . . . . . . . . . . . . . . 17

     2.1   Introduction . . . . . . . . . . . . . . . . . . . 17
     2.2   Basic Disk Operating System Functions  . . . . . . 29
     2.3   Queue and Process Descriptor Data Structures . . . 53
     2.4   Extended Disk Operating System Functions . . . . . 62
     2.5   Preparation of Page Relocatable Programs . . . . . 81
     2.6   Installation of Resident System Processes  . . . . 83

3.   MP/M Alteration Guide  . . . . . . . . . . . . . . . . . 85

     3.1   Introduction . . . . . . . . . . . . . . . . . . . 85
     3.2   Basic I/O System Entry Points  . . . . . . . . . . 96
     3.3   Extended I/O System Entry Points . . . . . . . .  102
     3.4   System File Components . . . . . . . . . . . . .  107
     3.5   System Generation  . . . . . . . . . . . . . . .  110
     3.6   MP/M Loader  . . . . . . . . . . . . . . . . . .  114
.BP
.SP 2
.CE
Appendix
.SP 2
.LI
     A.  Flag Assignments . . . . . . . . . . . . . . . . .  116

     B.  Process Priority Assignments . . . . . . . . . . .  117

     C.  BDOS Function Summary  . . . . . . . . . . . . . .  118

     D.  XDOS Function Summary  . . . . . . . . . . . . . .  119

     E.  Memory Segment Base Page Reserved Locations  . . .  120

     F.  Operation of MP/M on the Intel MDS-800 . . . . . .  121

     G.  Sample Page Relocatable Program  . . . . . . . . .  122

     H.  Sample Resident System Process . . . . . . . . . .  127

     I.  Sample XIOS  . . . . . . . . . . . . . . . . . . .  131

     J.  MP/M DDT Enhancements  . . . . . . . . . . . . . .  148

     K.  Page Relocatable (PRL) File Specification  . . . .  149

.AD
.bp
.sp 2
.ce
FOREWORD
.sp
.pp
This manual is intended as a guide for three different levels of
MP/M users.  Section 1 contains all the information required to
enable a person to operate applications programs running under the
MP/M Operating System.  Thus, the first section of this manual
should enable the casual user to operate the system with a minimum
amount of study and training.
.pp
The second section of this manual describes the MP/M system
organization including the structure of memory and system call
functions.  The intention is to provide the
necessary information required to write page relocatable programs and
resident system processes which operate under MP/M, and which use
the real-time multi-tasking, peripheral and disk I/O facilities of
the system.
.pp
The last section provides the information needed to tailor MP/M
to another computer system.  In particular, the hardware dependent
basic and extended I/O system entry points are
described.
Preparation of the MP/M loader using a CP/M 2.0 BIOS is also
covered.
.pp
The system generation procedure is also described in the last
section.  This procedure is of interest to all three levels of
MP/M users because it describes how to configure MP/M for a particular
applications environment.  This configuration includes the specificaion
of memory segmentation, number of consoles, and selection optional
resident system processes such as the printer spooler.
.br
