.MB +5
.MT -3
.LL 65
.PN 1
.HE MP/M User's Guide
.FT   (All Information Herein is Proprietary to Digital Research.)
.sp 2
1.  MP/M FEATURES AND FACILITIES
.sp
.pp
1.1  Introduction
.pp
The purpose of the MP/M multi-programming monitor control program is to
provide a microcomputer operating system which
supports multi-terminal access with multi-programming at each terminal.
.sp
.CE
OVERVIEW
.PP
The MP/M operating system is an upward compatible
version of CP/M 2.0 with a number of added facilities.  These added
facilities are contained in new logical sections of MP/M called the
extended I/O system and the extended
disk operating system.
In this manual the name XIOS will refer to the combined basic
and extended I/O system.  BDOS will refer to the standard CP/M 2.0
basic disk operating system functions and XDOS will refer to the
extended disk operating system.
As an upward compatible version, users can
easily make the transition from CP/M to the MP/M
operating system.
In fact, existing CP/M *.COM files can be run under MP/M,
providing that the program has been correctly written.  That is,
BDOS calls are made for I/O,
and the only direct BIOS calls made are for console and printer I/O.
There must also be at least 4 bytes of extra stack in the CP/M
*.COM program.
.SP
The following basic facilities are provided:
.LI

      o Multi-terminal support
      o Multi-Programming at each terminal
      o Support for bank switched memory and
          memory protection
      o Concurrency of I/O and CPU operations
      o Inter-process communication, mutual
          exclusion and synchronization
      o Ability to operate in sequential, polled
          or interrupt driven environments
      o System timing functions
      o Logical interrupt system utilizing flags
      o Selection of system options at system
          generation time
      o Dynamic system configuration at load time

The following optional facilities are provided:

      o Spooling list files to the printer
      o Scheduling programs to be run by date and time
      o Displaying complete system run-time status
      o Setting and reading of the date and time
.AD
.SP
.CE
HARDWARE ENVIRONMENT
.PP
The hardware environment for MP/M must include an 8080 or Z80 CPU,
a minimum of 32K bytes of memory, 1 to 16 consoles, 1 to 16
logical (or physical) disk drives
each containing up to eight megabytes, and a clock/timer interrupt.
.PP
The distributed form of the MP/M operating system is configured for
a polled I/O environment on the Intel MDS-800 with two consoles
and a real-time clock.  Multi-programming at two terminals is supported with this
configuration.  To improve the system performance and
capability the following incremental hardware additions can be utilized
by the operating system:
.SP
.IN +3
   a. Full Interrupt System
.BR
   b. Banked Memory
.BR
   c. Additional Consoles
.QI
.AD
.SP
.CE
MEMORY SIZE
.PP
The MP/M operating system requires less than
15K bytes of memory when configured for
two consoles and eight memory segments on the Intel MDS-800.
Each additional console
requires 256 bytes.
.PP
Optional resident system processes can be specified at system
generation which require varying amounts of memory.
.SP
.CE
PERFORMANCE
.PP
When MP/M is configured for a single console and is executing
a single process, its speed approximates that of CP/M.
In environments where either multiple processes and/or users are
running, the speed of each individual process is degraded in
proportion to the amount of I/O and compute resources required.
A process
which performs a large amount of I/O in proportion to computing
exhibits only minor speed degradation.  This also applies to a process
that performs a large amount of computing, but is running concurrently
with other processes that are largely I/O bound.  On the other hand,
significant speed degradation occurs in environments in which
more than one compute bound process is running.
.bp
.sp
.pp
1.2  Functional Description of MP/M
.pp
The MP/M Operating System is based on a real-time multi-tasking
nucleus.  This nucleus provides process dispatching, queue management,
flag management, memory management and system timing functions.
.pp
MP/M is a priority driven system.  This means that the highest priority
ready process is given the CPU resource.  The operation of determining
the highest priority ready process and then giving it the CPU is called
dispatching.  Each process in the system has a process descriptor.  The
purpose of the process descriptor is to provide a data structure which
contains all the information the system needs to know about a process.
This information is used during dispatching to save
the state of the currently running process, to determine which process is
to be run, and then to restore that processes state.
Process dispatching is performed at each system call, at each
interrupt, and at each tick of the system clock.
Processes with the same priority are "round-robin" scheduled. That is,
they are given equal slices of CPU time.
.pp
Queues perform several critical functions in a real-time multi-tasking
environment.  They can be used for the communication of messages
between processes, to synchronize processes, and for mutual exclusion.
As the name "queue" implies, they provide a first in first out list of
messages, and as implemented in MP/M, a list of processes waiting for
messages.
.pp
The flag management provided by MP/M is used to synchronize processes
by signaling a significant event.  Flags provide
a logical interrupt system for MP/M which is independent of the
physical interrupt system.  Flags are used to signal interrupts,
mapping an arbitrary physical interrupt environment into a regular
structure.
.pp
MP/M manages memory in pre-defined memory segments.
Up to eight memory segments of 48K can be managed by MP/M.
This management
of memory is consistent with hardware environments where memory
is banked and/or protected in fixed segments.
.pp
System timing functions provide time of day, the capability to
schedule programs to be loaded from disk and executed, and the
ability to delay the execution of a process for a specified period
of time.
.bp
.sp
.pp
1.3  Console Commands
.PP
The purpose of this section is to describe the console commands
which make up the operator interface to the MP/M operating
system.  It is important to note from the outset that there are
no system defined or built-in commands.  That is, the system has no
reserved or special commands.  All commands in the system are
provided by resident system processes specified during system
generation and programs residing on disk in either the CP/M
*.COM file format or in the MP/M *.PRL (page relocatable) file format.
.pp
When MP/M is loaded from disk a configuration table and memory segment map
are displayed on console #0.
When the loading is complete each of the 1 to 16
configured consoles is a system or master console.
Additional slave consoles (maximum total of slave and master consoles
is 16) can be accessed using XDOS system calls.
.pp
After loading, the following message is displayed on each console:
.sp
.ti +3
MP/M
.ti +3
xA>
.pp
The 'x' shown in the prompt is the user code.  At cold start an
association is made between the user code and console number.  The
initial user code is equal to the console number.  For
example, console #0 is initialized to user #0 and the following
prompt is displayed on console #0:
.sp
.ti +3
0A>
.pp
The default user code can then be changed to any desired user code
with the USER command (see USER in section 1.4).
All users have access to files with a user code of 0.  Thus, system
files and programs should have a user code of 0.
Caution must be used when operating under a user
code of 0 since all its files can be accessed while
operating under any other user code.
In general, user code 0 should be reserved for
files which are accessed by all users.  In the event that a
file  with  the same name is present under user code 0  and
another  user code,  the first file found in the  directory
will be accessed.
.pp
The 'A' in the prompt is the default (currently logged) disk for the
console.  This can be changed individually at any console by typing
in a disk drive name (A,B,C,...,or P) followed by a colon (:) when
the prompt has been received.
Since  there  are no built-in commands,  the  default  disk
specified  must contain the desired command files (such  as
DIR,  REN,  ERA etc.), or each command must be preceeded by
an "A:".
.SP
.CE
RUNNING A PROGRAM
.PP
A program is run by typing in the program name followed by a carriage return, <cr>.  Some programs obtain parameters on the same line
following the program name.  Characters on the line following the
program name constitute what is called the command tail.  The command
tail is copied into location 0080H (relative to the base of the memory
segment in which the program resides) and converted to upper case by the Command Line Interpreter
(CLI).  The CLI also parses the command tail producing two file control
blocks at 005CH and 006CH respectively.
.PP
The programs which are provided with MP/M are described in sections
1.4 and 1.5.
.SP
.CE
ABORTING A PROGRAM
.PP
A program may be aborted by typing a control C (^C) at the console.
The affect of the ^C is to terminate the program which currently
owns the console.  Thus, a detached program cannot be aborted with
a ^C.  A detached program must first be attached and then aborted.
A running program may also be aborted using the ABORT command (see
ABORT in section 1.5).
.SP
.CE
RUNNING A RESIDENT SYSTEM PROCESS
.PP
At the operator interface there is no difference between running
a program from disk and running a resident system process.  The actual
difference is that resident system processes do not need to be
loaded from disk because they are loaded by the MP/M loader when
a system cold start is performed and remain resident.
.SP
.CE
DETACHING FROM A PROGRAM
.PP
There are two methods for detaching from a running program.  The first
is to type a control D (^D) at the console.  The second method is for
a program to make an XDOS detach call.
.PP
The restriction on the former method, typing ^D, is that the running
program must be performing a check console status to observe the
detach request.
A check console status is automatically performed each time
a user program makes a BDOS disk function call.
.SP
.CE
ATTACHING TO A DETACHED PROGRAM
.PP
A program which is detached from a console, that is it does not own a
console, may be attached to a console by typing 'ATTACH' followed by
the program name.  A program may only be attached to the console from
which it was detached.
If the terminal message process (TMP)
has ownership of the console and the user enters a ^D, the
next highest priority ready process which is waiting for the console
begins running.
.SP
.CE
LINE EDITING AND OUTPUT CONTROL
.PP
The Terminal Message Process (TMP) allows certain line editing
functions while typing in command lines:
.LI

	rubout	Delete the last character typed at the console,
	        removes and echoes the last character

	ctl-C	MP/M abort program. Terminate running process.

	ctl-D   MP/M detach console.

	ctl-E   Physical end of line.

	ctl-H   Delete the last character typed at the console,
	        backspaces one character position.

	ctl-J   (line feed) terminate current input.

	ctl-M   (carriage return) terminates input.

	ctl-R	Retype current command line: types a "clean line"
		following character deletion with rubouts.

	ctl-U   Remove current line after new line.

	ctl-X	Delete the entire line typed at the console,
	        backspaces to the beginning of the current line

	ctl-Z	End input from the console.

   The control functions ctl-P, ctl-Q and ctl-S affect console
   output as shown below.

	ctl-P	Copy all subsequent console output to the list
		device.  Output is sent to both the list device
		and the console device until the next ctl-P
		is typed.  If the list device is not available
		a 'Printer busy' message is displayed on the
		console.

	ctl-Q	Obtain ownership of the printer mutual exclusion
		message.  Obtaining the printer using this command
		will ensure that the MP/M spooler, PIP, and other
		ctl-P or ctl-Q commands entered from other
		consoles will not be allowed access to the
		printer.  The printer is "owned" by the TMP until
		another ctl-P or ctl-Q is entered, releasing the
		printer.  The ctl-P should be used when a program
		(such as a CP/M *.COM file) is executed that does
		not obtain the printer mutual exclusion message
		prior to accessing the printer.  If the list
		device is not available a 'Printer busy' message
		is displayed on the console.

	ctl-S	Stop the console output temporarily.  Program
		execution and output continue when the next
		character is typed at the console (e.g., another
		ctl-S).  This feature is used to stop output on
		high speed consoles, such as CRT's, in order to
		view a segment of output before continuing.

.AD
.bp
.sp
.pp
1.4  Commonly Used System Programs
.PP
The commonly used system programs (CUSPs) or transient commands, as
they are called in CP/M, are loaded from the currently logged
disk and executed in a relocatable memory segment if their type is
PRL or in an absolute TPA if their type is COM.
.PP
This section contains a brief description of the CUSPs.  Operation of
many of the CUSPs is identical to that under CP/M.  In these cases
the commands are marked with an asterisk '*' and
the reader is referred to the Digital Research document titled "An
Introduction to CP/M Features and Facilities" for a complete
description of the CUSP.
.SP
.CE
GET/SET USER CODE
.PP
The USER command is used to display the current user code as well as
to set the user code value.
Entering the command USER followed by a <cr> will display the
current user code.
Note that the user code is already displayed in the prompt.
.sp
.ti +3
1A>user
.ti +3
user = 1
.pp
Entering the command USER followed by a space, a user code and then
a <cr> will set the user code to the specified user code.  Legal user
codes are in the range 0 to 15.
.sp
.ti +3
1A>user 3
.ti +3
user = 3
.ti +3
3A>
.sp
.CE
CONSOLE
.pp
The CONSOLE command is used to determine the console number at
which the command is entered.  The console number is sometimes of
interest when examining the system status to determine the processes
which are detached from consoles.
.sp
.ti +3
1A>console
.ti +3
Console = 0
.sp
.ce
DISK RESET
.pp
The  DSKRESET  (disk reset) command is used to  enable  the
operator to change disks.   If no parameter is entered  all
the  drives are reset.   Specific drives to be reset may be
included as parameters.
.sp
.ti +3
1A>DSKRESET
.sp
.ti +3
1A>DSKRESET  B:,E:
.pp
If  there are any open files on the drive(s) to  be  reset,
the  disk  reset is denied and the cause of the disk  reset
failure is shown:
.sp
.ti +3
1A>DSKRESET B:
.sp
.ti +3
Disk reset denied, Drive B: Console 0 Program Ed
.pp
The reason that disk reset is treated so carefully is  that
files left open (e.g. in the process of being written) will
lose their updated information if they are not closed prior
to a disk reset.
.sp
.ce
ERASE FILE *
.PP
The ERA (erase) command removes specified files having the current
user code.
If no files can be found on the selected diskette which satisfy the
erase request, then the message "No file" is displayed at the console.
.sp
An attempt to erase all files,
.sp
.ti +3
2B>ERA *.*
.sp
will produce the following response from ERA:
.sp
.ti +3
Confirm delete all user files (Y/N)?
.pp
A  second  form  of the erase command  (ERAQ)  enables  the
operator  to  selectively  delete  files  that  match   the
specified filename reference.  For example:
.sp
.ti +3
0A>ERAQ *.LST
.sp
.ti +3
A:XIOS    LST? y
.ti +3
A:MYFILE  LST? n
.SP
.CE
TYPE A FILE *
.PP
The TYPE command displays the contents of the specified ASCII source
file on the console device.  The TYPE command expands tabs (ctl-I
characters), assuming tab positions are set at every eighth column.
.pp
The  TYPE  command has a pause mode which is  specified  by
entering  a  'P' followed by two decimal digits  after  the
filename. For example:
.sp
.ti +3
0A>TYPE DUMP.ASM P23
.sp
The  specified number of lines will be displayed  and  then
TYPE will pause until a <cr> is entered.
.pp
The  TYPE  program is small and relatively slow because  it
buffers only one sector at a time.   The larger PIP program
can be used for faster displays in the following manner:
.sp
.ti +3
0A>PIP CON:=MYFILE.TEX
.SP
.CE
FILE DIRECTORY *
.PP
The DIR (directory) command causes the names of files on the specified or logged-in
disk to be listed on the console device.  If no files can be found
on the selected diskette which satisfy the directory request, then
the message "Not found" is typed at the console.
.pp
The  DIR  command can include files which have  the  system
attribute set.   This is done by using the 'S' option.  For
example:
.sp
.ti +3
0A>DIR *.COM S
.SP
.CE
RENAME FILE *
.PP
The REN (rename) command allows the user to change the name of files
on disk.
If  the  destination filename exists the operator is  given
the option of deleting the current destination file  before
renaming the source file.
.SP
.CE
TEXT EDITOR *
.PP
The ED (editor) command allows the user to edit ASCII text files.
.SP
.CE
PERIPHERAL INTERCHANGE PROGRAM *
.PP
The PIP (peripheral interchange program) command allows the user to
perform disk file and peripheral transfer operations.
See the Digital Research document titled "CP/M 2.0 User's Guide for
CP/M 1.4 Owners" for a detailed description of new PIP operations.
.SP
.CE
ASSEMBLER *
.PP
The ASM (assembler) command allows the user to assemble the specified
program on disk.
.SP
.CE
SUBMIT *
.PP
The SUBMIT command allows the user to submit a file of commands for
batch processing.
.SP
.CE
STATUS *
.PP
The STAT (status) command provides general statistical information
about the file storage.
See the Digital Research document titled "CP/M 2.0 User's Guide for
CP/M 1.4 Owners" for a detailed description of new STAT operations.
.SP
.CE
DUMP *
.PP
The DUMP command types the contents of the specified disk file on the
console in hexadecimal form.
.SP
.CE
LOAD *
.PP
The LOAD command reads the specified disk file of type HEX and
produces a memory image file of type COM which can subsequently
be executed.
.SP
.CE
GENMOD
.PP
The GENMOD command accepts a file which contains two concatenated
files of type HEX which are offset from each other by 0100H bytes,
and produces a file of type PRL (page relocatable).
The form of the GENMOD command is as follows:
.sp
.ti +3
1A>genmod b:file.hex b:file.prl $1000
.sp
The first parameter is the file which contains two concatenated files
of type HEX.  The second parameter is the name of the destination file
of type PRL.  The optional third parameter is a specification of additional memory required by the program beyond the explicit code
space.  The form of the third parameter is a '$' followed by four
hex ASCII digits.
For example, if the program has been written to use all of 'available'
memory for buffers, specification of the third parameter will ensure
a minimum buffer allocation.
.SP
.CE
GENHEX
.PP
The  GENHEX  command is used to produce a file of type  HEX
from  a  file of type COM.   This is useful to be  able  to
generate  HEX files for GENMOD input.   The GENHEX  command
has two parameters,  the first is the COM file name and the
second is the offset for the HEX file.  For example:
.sp
.ti +3
0A>GENHEX PROG.COM 100
.SP
.CE
PRLCOM
.PP
The  PRLCOM command accepts a file of PRL type and produces
a file of COM type.  If the destination COM file exists, a
query is made to determine if the file should be deleted
before continuing.
.sp
.ti +3
0A>prlcom b:program.prl a:program.com
.SP
.CE
DYNAMIC DEBUGGING TOOL *
.PP
The DDT (dynamic debugging tool) command loads and executes the
MP/M debugger.
In systems with banked memory multiple DDT programs can be running
concurrently in absolute TPAs.  A PRL (relocatable)
version of DDT is also provided
which enables mutiple DDTs to run in a non-banked system.  The
name of the relocatable DDT is RDT.
.pp
MP/M DDT enhancements are described in Appendix J.
.BP
.sp
.pp
1.5  Standard Resident System Processes
.PP
The standard resident system processes (RSPs) are new programs
specifically
designed to facilitate use of the MP/M operating system.  The RSPs
may either be present on disk as files of the PRL type, or they
may be resident system processes.  Resident system processes are
selected at the time of system generation.
.SP
.CE
SYSTEM STATUS
.PP
The MPMSTAT command allows the user to display the run-time status
of the MP/M operating system.  MPMSTAT is invoked by typing 'MPMSTAT'
followed by a <cr>.  A sample MPMSTAT output is shown below:
.LI

	****** MP/M Status Display ******
	
	Top of memory = FFFFH
	Number of consoles = 02
	Debugger breakpoint restart # = 06
	Stack is swapped on BDOS calls
	Z80 complementary registers managed by dispatcher
	Ready Process(es):
	  MPMSTAT  Idle
	Process(es) DQing:
	 [Sched   ] Sched
	 [ATTACH  ] ATTACH
	 [CliQ    ] cli 
	Process(es) NQing:
	Delayed Process(es):
	Polling Process(es):
	  PIP
	Process(es) Flag Waiting:
	  01 -  Tick
	  02 -  Clock
	Flag(s) Set:
	  03
	Queue(s):
          MPMSTAT   Sched     CliQ      ATTACH    MXParse
          MXList   [Tmp0    ] MXDisk
	Process(es) Attached to Consoles:
	  [0] - MPMSTAT
	  [1] - PIP
	Process(es) Waiting for Consoles:
	  [0] - TMP0     DIR
	  [1] - TMP1
	Memory Allocation:
	  Base = 0000H  Size = 4000H  Allocated to PIP      [1]
	  Base = 4000H  Size = 2000H  * Free *
	  Base = 6000H  Size = 1100H  Allocated to DIR      [0]
.AD
.bp
.pp
.cp 4
The MP/M status display is intepreted as follows:
.in +6
.pp
Ready Process(es):  The ready processes are those processes which
are ready to run and are waiting for the CPU.  The list of ready
processes is ordered by the priority of the processes and includes
the console number at which the process was initiated.  The highest
priority ready process is the running process.
.pp
Process(es) DQing:  The processes DQing are those processes which are
waiting for messages to be written to the specified queue.  The queue
name is in brackets followed by the names of processes, in priority
order, which have executed read queue operations on the queue.
.pp
Process(es) NQing:  The processes NQing are those processes which are
waiting for an available buffer to write a message to the specified
queue.  The queue name is in brackets followed by the names of the
processes, in priority order, which are waiting for buffers.
.pp
Delayed Process(es):  The delayed processes are those which are
delaying for a specified number of ticks of the system time unit.
.pp
Polling Process(es):  The polling processes are those which
are polling a specified I/O device for a device ready status.
.pp
Process(es) Flag Waiting:  The processes flag waiting are listed by
flag number and process name.
.pp
Flag(s) Set:  The flags which are set are displayed.
.pp
Queue(s):  All the queues in the system are listed by queue name.
Queue names which are all in capital letters are accessible by
command line interpreter
input.  For example, the SPOOL queue can be sent a message to spool
a file by entering 'SPOOL' followed by a file name.
Processes DQing from queues which have a name that matches the
process name are given the console resource when they receive a
message.
Queue names that begin with 'MX' are called mutual exclusion queues.
The display of a mutual exclusion queue includes the name of the process,
if any, which has the mutual exclusion message.
.pp
Process(es) Attached to Consoles:  The process attached to each console
is listed by console number and process name.
.pp
Process(es) Waiting for Consoles:  The processes waiting for each
console are listed by console number and process name in priority
order.  They are processes
which have detached from the console and are then waiting for the
console before they can continue execution.
.pp
Memory Allocation:  The memory allocation map shows the base, size, bank,
and allocation of each memory segment.  Segments which are not
allocated are shown as '* Free *', while allocated segments are
identified by process name and the console in brackets associated with
the process.
Memory segments which are set as pre-allocated during system generation
by specifying an attribute of 0FFH are shown as '* Reserved *'.
.qi
.sp
.CE
SPOOLER
.PP
The SPOOL command allows the user to spool ASCII text files to the
list device.  Multiple file names may be specified in the command
tail.  The spooler expands tabs (ctl-I characters), assuming tab
positions are set at every eighth column.
.PP
The spooler queue can be purged at any time by using the STOPSPLR
command.
.PP
An example of the SPOOL command is shown below:
.sp
.ti +3
1A>SPOOL LOAD.LST,LETTER.PRN
.pp
The non-resident version of the spooler (SPOOL.PRL) differs
in its operation from the SPOOL.RSP as follows: it uses all
of  the memory available in the memory segment in which  it
is  running  for  buffer  space;   it  displays  a  message
indicating  its status and then detaches from the  console;
it  may be aborted from a console other than the  initiator
only by specifying the console number of the initiator as a
parameter of the STOPSPLR command.
.sp
.ti +3
3B>STOPSPLR 2
.SP
.CE
DATE AND TIME
.PP
The TOD (time of day) command allows the user to read and set the
date and time.  Entering 'TOD' followed by a <cr> will cause the
current date and time to be displayed on the console. Entering
'TOD' followed by a date and time will set the date and time
when a <cr> is entered following the prompt to strike a key.
Each of these TOD commands is illustrated below:
.sp
.li
   1A>TOD <cr>

   Wed 02/06/80 09:15:37

   -or-

   1A>TOD 2/9/80 10:30:00
   Strike key to set time
   Sat 02/09/80 10:30:00
.AD
.pp
Entering 'TOD P' will cause the current time and date to be
continuously displayed until a key is struck at the console.
.SP
.CE
SCHEDULER
.PP
The SCHED (scheduler) command allows the user to schedule a program
for execution.  Entering 'SCHED' followed by a date, time and command
line will cause the command line to be executed when the specified
date and time is reached.
.PP
In the example shown below, the program 'SAMPLE' will be loaded from
disk and executed on February 8, 1980 at 10:30 PM.
Note that only hours and minutes are specified, not seconds.  Programs
are scheduled to the nearest minute.
.sp
.ti +3
1A>SCHED 2/8/79 22:30 SAMPLE
.SP
.CE
ABORT
.PP
The  ABORT  command  allows the user  to  abort  a  running
program.   The  program  to  be  aborted is  entered  as  a
parameter in the ABORT command.
.sp
.ti +3
1A>ABORT RDT
.pp
A  program  initiated  from another  console  may  only  be
aborted by including its console number as
a parameter of the ABORT command.
.sp
.ti +3
3B>ABORT RDT 1
.sp
.br
