.MB +5
.MT -3
.LL 65
.PN 53
.HE MP/M User's Guide
.FT   (All Information Herein is Proprietary to Digital Research.)
.sp
.pp
2.3  Queue and Process Descriptor Data Structures
.pp
This section contains a description of the queue and process
descriptor data structures used by the MP/M Extended Disk Operating
System (XDOS).
.sp
.ce
QUEUE DATA STRUCTURES
.pp
A queue is a first in first out (FIFO) mechanism which has been
implemented in MP/M to provide several essential functions in a
multi-tasking environment.  Queues can be used for the communication
of messages between processes, to synchronize processes, and to provide
mutual exclusion.
.pp
MP/M has been designed to simplify queue management for both user
and system processes.  In fact, queues are treated in a manner
similar to disk files.  Queues can be created, opened, written to,
read from, and deleted.
.pp
A few illustrations should suffice to describe applications
for queues:
.pp
COMMUNICATION:
.pp
A queue can be used for communication to provide a FIFO list of
messages produced by a producer for consumption by a consumer.
For example, consider a data logging application where
data is continuously received via a serial communication link and is to
be written to a disk file.  This would be a difficult application
for a sequential operating system such as CP/M because arriving
serial data would be lost while buffers were being written to
disk.  Under MP/M a queue could be used by the producer to send
blocks of received serial data (or simply buffer pointers) to a consumer
which would write the blocks on disk.  MP/M supports concurrency of
these operations, allowing the producer to quickly write a buffer to
the queue and then resume monitoring the serial input.
.pp
SYNCHRONIZATION:
.pp
When a process attempts to read a message at a queue and there are
no messages posted at the queue, the process is placed in a priority
ordered list of processes waiting for messages at the queue.  The
process will remain in that state until a message arrives.  Thus
synchronization of processes can be achieved, allowing the waiting
(DQing) process to continue execution when a message is sent to
the queue.
.cp 20
.pp
MUTUAL EXCLUSION:
.pp
A queue can also be used for mutual exclusion.
Mutual exclusion messages generally have a length of zero.  A good
example of mutual exclusion is that which is used by MP/M to
control access to the printer.  A queue is created (MXList)
and sent one message.  When the printer is to be used by the spooler
or by entering a control-P (^P) at the console an attempt is
made to read the message from the list mutual exclusion queue.
This attempt is made using the MP/M conditional read queue function.
If the message is available, that is it has not been consumed by
some other process, it is read and the printer is used.  When finished
with the printer,
the message is written back to the list mutual exclusion queue.
If the message is not available the user who entered the ^P
receives a message indicating that the printer is busy.
In the case of the spooler, it waits until the printer is available
before continuing.
.sp
.ce
QUEUE DATA STRUCTURES
.SP
.pp
The queue data structures include the queue control block and the user
queue control block.  There are two types of queue control blocks,
circular or linked.  The type of queue control block used depends upon
the message size.  Message sizes of 0 to 2 bytes use circular queues
while message sizes of 3 or more bytes use linked queues.
.sp
.ce
CIRCULAR QUEUES
.pp
The following example illustrates how to setup a queue control block
for a circular queue with 80 messages of a one byte length.
Each example in this section will be shown both in PL/M and assembly
language.
.cp 17
.LI

   PL/M:

	DECLARE CIRCULAR$QUEUE STRUCTURE (
	  QL ADDRESS,
	  NAME(8) BYTE,
	  MSGLEN ADDRESS,
	  NMBMSGS ADDRESS,
	  DQPH ADDRESS,
	  NQPH ADDRESS,
	  MSG$IN ADDRESS,
	  MSG$OUT ADDRESS,
	  MSG$CNT ADDRESS,
	  BUFFER (80) BYTE        )
	  INITIAL (0,'CIRCQUE ',1,80);
.ad
.cp 14
.li

   Assembly Language:

	CRCQUE:
		DS	2	; QL
		DB	'CIRCQUE ' ; NAME
		DW	1	; MSGLEN
		DW	80	; NMBMSGS
		DS	2	; DQPH
		DS	2	; NQPH
		DS	2	; MSGIN
		DS	2	; MSGOUT
		DS	2	; MSGCNT
	BUFFER:	DS	80	; BUFFER

.AD
The elements of the circular queue shown above are defined
as follows:
.LI

	QL      = 2 byte link, set by system
	NAME    = 8 ASCII character queue name,
	            set by user
	MSGLEN  = 2 bytes, length of message,
	            set by user
	NMBMSGS = 2 bytes, number of messages,
	            set by user
	DQPH    = 2 bytes, DQ process head,
	            set by system
	NQPH    = 2 bytes, NQ process head,
	            set by system
	MSG$IN  = 2 bytes, pointer to next
	            message in, set by system
	MSG$OUT = 2 bytes, pointer to next
	            message out, set by system
	MSG$CNT = 2 bytes, number of messages
	            in the queue, set by system
	BUFFER  = n bytes, where n is equal to
	            the message length times the
	            number of messages, space
	            allocated by user, set by system

                    Note: Mutual exclusion queues require
                    a two byte buffer for the owner process
                    descriptor address.

Queue Overhead  = 24 bytes


.AD
.sp
.ce
LINKED QUEUES
.pp
The following example illustrates how to setup a queue control block
for a linked queue containing 4 messages, each 33 bytes in length:
.cp 16
.LI

   PL/M:

	DECLARE LINKED$QUEUE STRUCTURE (
	  QL ADDRESS,
	  NAME (8) BYTE,
	  MSGLEN ADDRESS,
	  NMBMSGS ADDRESS,
	  DQPH ADDRESS,
	  NQPH ADDRESS,
	  MH ADDRESS,
	  MT ADDRESS,
	  BH ADDRESS,
	  BUFFER (140) BYTE   )
	  INITIAL (0,'LNKQUE  ',33,4);
.ad
.cp 21
.li

   Assembly Language:

	LNKQUE:
		DS	2	; QL
		DB	'LNKQUE  ' ; NAME
		DW	33	; MSGLEN
		DW	4	; NMBMSGS
		DS	2	; DQPH
		DS	2	; NQPH
		DS	2	; MH
		DS	2	; MT
		DS	2	; BH
	BUFFER:	DS	2	; MSG #1 LINK
		DS	33	; MSG #1 DATA
		DS	2	; MSG #2 LINK
		DS	33	; MSG #2 DATA
		DS	2	; MSG #3 LINK
		DS	33	; MSG #3 DATA
		DS	2	; MSG #4 LINK
		DS	33	; MSG #4 DATA

.AD
The elements of the linked queue shown above are defined
as follows:
.LI

	QL      = 2 byte link, set by system
	NAME    = 8 ASCII character queue name,
	            set by user
	MSGLEN  = 2 bytes, length of message,
	            set by user
	NMBMSGS = 2 bytes, number of messages,
	            set by user
	DQPH    = 2 bytes, DQ process head,
	            set by system
	NQPH    = 2 bytes, NQ process head,
	            set by system
	MH      = 2 bytes, message head,
	            set by system
	MT      = 2 bytes, message tail,
	            set by system
	BH      = 2 bytes, buffer head,
	            set by system
	BUFFER  = n bytes where n is equal to
	            the message length plus two,
	            times the number of messages,
	            space allocated by the user,
	            set by the system

.AD
.sp
.ce
USER QUEUE CONTROL BLOCK
.pp
The user queue control block data structure is used to provide
read/write access to queues in much the same manner that a file
control block provides access to a disk file.  Queues are "opened",
an operation which fills in the actual queue control block address,
and then can be read from or written to.
.pp
If the actual queue address is known it can be filled in the
pointer field of the user queue control block, the 8 byte name
field can be omitted, and an open operation is not required in
order to access the queue.
.pp
The following example illustrates a user queue control block:
.cp 9
.LI

   PL/M:

	DECLARE USER$QUEUE$CONTROL$BLOCK STRUCTURE (
	  POINTER ADDRESS,
	  MSGADR ADDRESS,
	  NAME (8) BYTE    )
          INITIAL (0,.BUFFER,'SPOOL   ');

	DECLARE BUFFER (33) BYTE;
.ad
.cp 11
.li

   Assembly Language:

	UQCB:
		DS	2	; POINTER
		DW	BUFFER	; MSGADR
		DB	'SPOOL   ' ; NAME

	BUFFER:
		DS	33	; BUFFER

.AD
.cp 14
.pp
The elements of the user queue control block shown above are defined
as follows:
.LI

	POINTER = 2 bytes, set by system to address of
	            actual queue during an open queue
	            operation, or set by the user if the
	            actual queue address is known
	MSGADR  = 2 bytes, address of user buffer,
	            set by user
	NAME    = 8 bytes, ASCII queue name,
	            set by user, may be omitted if the
	            pointer field is set by the user

.AD
.sp
.ce
QUEUE NAMING CONVENTIONS
.pp
The  following conventions should be used in the naming  of
queues.  Queues which are to be directly written to by the
Terminal   Message  Process  (TMP)  via  the  Command  Line
Interpreter (CLI) must have an upper case ASCII name.  Thus
when an operator enters the queue name followed by a
command  tail at a console,  the command tail is written to
the queue.
.pp
In  order  to  make a queue inaccessible by  a  user  at  a
console it must contain at least one lower case character.
Mutual  exclusion  queues should be named upper  case  'MX'
followed  by  1 to 6 additional  ASCII  characters.  These
queues  are treated specially in that they must have a  two
byte buffer in which MP/M places the address of the process
descriptor owning the mutual exclusion message.
.cp 35
.sp
.ce
PROCESS DESCRIPTOR
.pp
Each process in the MP/M system has a process descriptor which
defines all the characteristics of the process.  The following
example illustrates the process descriptor:
.cp 28
.LI

   PL/M:

	DECLARE CNS$HNDLR STRUCTURE (
	  PL ADDRESS,
	  STATUS BYTE,
	  PRIORITY BYTE,
	  STKPTR ADDRESS,
	  NAME (8) BYTE,
	  CONSOLE BYTE,
	  MEMSEG BYTE,
	  B ADDRESS,
	  THREAD ADDRESS,
	  DISK$SET$DMA ADDRESS,
	  DISK$SLCT BYTE,
	  DCNT ADDRESS,
	  SEARCHL BYTE,
	  SEARCHA ADDRESS,
	  DRVACT ADDRESS,
	  REGISTERS (20) BYTE,
	  SCRATCH (2) BYTE )
	  INITIAL (0,0,200,.CNS$STK(19),
	           'CNS     ',1,0FFH);

	DECLARE CNS$STK (20) ADDRESS INITIAL (
	  0C7C7H,0C7C7H,0C7C7H,0C7C7H,0C7C7H,0C7C7H,
	  0C7C7H,0C7C7H,0C7C7H,0C7C7H,0C7C7H,0C7C7H,
	  0C7C7H,0C7C7H,0C7C7H,0C7C7H,0C7C7H,0C7C7H,
	  0C7C7H,STRT$CNS);
.ad
.cp 28
.li

   Assembly Language:

	CNSHND:
		DW	0	; PL
		DB	0	; STATUS
		DB	200	; PRIORITY
		DW	CNSTK+38 ;STKPTR
		DB	'CNS     ' ; NAME
		DB	0	; CONSOLE
		DB	0FFH	; MEMSEG (FF = resident)
		DS	2	; B
		DS	2	; THREAD
		DS	2	; DISK SET DMA
		DS	1	; DISK SLCT
		DS	2	; DCNT
		DS	1	; SEARCHL
		DS	2	; SEARCHA
		DS	2	; DRVACT
		DS	20	; REGISTERS
		DS	2	; SCRATCH

	CNSTK:
		DW	0C7C7H,0C7C7H,0C7C7H,0C7C7H
		DW	0C7C7H,0C7C7H,0C7C7H,0C7C7H
		DW	0C7C7H,0C7C7H,0C7C7H,0C7C7H
		DW	0C7C7H,0C7C7H,0C7C7H,0C7C7H
		DW	0C7C7H,0C7C7H,0C7C7H
		DW	CNSPR	; CNSTK+38 = PROCEDURE ADR
.AD
.pp
The elements of the process descriptor shown above are defined
as follows:
.LI

	PL            = 2 byte link field, initially set by
	                  user to address of next process
	                  descriptor, or zero if no more
	STATUS        = 1 byte, process status, set by system
	PRIORITY      = 1 byte, process priority, set by user
	STKPTR        = 2 bytes, stack pointer, initially set
	                  by user
	NAME          = 8 bytes, ASCII process name, set by user

                      The  high  order  bit of each byte of the
                      process  name  is  reserved for use by the
                      system.  The high order bit of the first
                      byte (identified as NAME(0)') "on" indicates
                      that the process  is  performing direct
                      console BIOS calls and  that  MP/M is to
                      ignore all control characters.  It is also
                      used  to suppress the normal console status
                      check done when  BDOS  disk functions are
                      invoked.  This bit may be set by the user.

	CONSOLE       = 1 byte, console to be used by process,
	                  set by user
	MEMSEG        = 1 byte, memory segment table index
	B             = 2 bytes, system scatch area
	THREAD        = 2 bytes, process list thread, set
	                  by system
	DISK$SET$DMA  = 2 bytes, default DMA address, set by user
	DISK$SLCT     = 1 byte, default disk/user code
	DCNT          = 2 bytes, system scratch byte
	SEARCHL       = 1 byte, system scratch byte
	SEARCHA       = 2 bytes, system scratch bytes
	DRVACT        = 2 bytes, 16 bit vector of drives being
	                accessed by the process
	REGISTERS     =20 bytes, 8080 / Z80 register save area
	SCRATCH       = 2 bytes, system scratch bytes
	
.AD
.sp
.ce
PROCESS NAMING CONVENTIONS
.pp
The  following conventions should be used in the naming  of
processes.  Processes which wait on queues that are to be
sent command tails from the TMPs are given the console
resource if their name matches that of the queue which they
are reading.
Processes  which  are to be protected from abortion  by  an
operator  using  the ABORT command must have at  least  one
lower case character in the process name.
.br
