.bp odd
.he
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.hm 2
.fm 2
.ft All Information Presented Here is Proprietary to Digital Research
.ce 2
.sh
Section 7
.sp
.sh
Cold Boot Automatic Command Execution
.tc 7  Cold Boot Automatic Command Execution
.ix Cold Boot Automatic Command Execution
.sp 3
.fi
.sh
7.1  Overview
.tc    7.1  Overview
.pp
The Cold Boot Automatic Command Execution feature of CP/M-8000 allows you to 
configure CP/M-8000 so that the CCP will automatically execute a predetermined 
command line on cold boot.  This feature can be used to start up turn-key 
systems, or to perform other desired operations.
.ix turn-key systems
.sp
.sh
7.2  Setting up Cold Boot Automatic Command Execution
.ix CBACE feature
.pp
The CBACE feature uses two global symbols: _autost, and _usercmd.  These are 
both defined in the CCP, which uses them on cold boot
to determine whether this feature is enabled.  If you want to have a CCP
command automatically executed on cold boot, you should include code in your
BIOS's _init routine (which is called at cold boot) to do the following:
.ix _autost
.ix _usercmd
.ix _init routine
.sp 2
.in 8
.ti -3
1) The byte at _autost must be set to the value 01H.
.sp
.ti -3
2) The command line to be executed must be placed in memory at _usercmd and
subsequent locations.  The command must be terminated with a NULL (00H) byte,
and may not exceed 128 bytes in length.  All alphabetic characters in the
command line should be upper-case.
.sp
.in 0
.pp
Once you write a BIOS that performs these two functions, you can 
build it into a CPM.SYS file as described in Section 2.  This system, when
booted, will execute the command you have built into it.
.ix CPM.SYS file
.sp 2
.ce
End of Section 7
.bp
.he CP/M-8000 System Guide                            End of Section 7
.sp 30
.bp
.nx eight
