.bp odd
.he
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.hm 2
.fm 2
.ft All Information Presented Here is Proprietary to Digital Research
.ce 2
.sh
Section 8
.sp
.sh
The PUTBOOT Utility
.tc 8  The PUTBOOT Utility
.sp 3
=============== WE OR SOMEBODY NEEDS TO WRITE THIS UTILITY =============
.fi
.sh
8.1  PUTBOOT Operation
.tc    PUTBOOT Operation      
.ix PUTBOOT utility
.pp 5
The PUTBOOT utility is used to copy information (usually a bootstrap loader
system) onto the system tracks of a disk.  Although PUTBOOT can copy any 
file to the system tracks, usually the file being written is a program 
(the bootstrap system).  
.sp 2
.sh
8.2  Invoking PUTBOOT
.tc    8.2  Invoking PUTBOOT
.he CP/M-8000 System Guide                            Invoking PUTBOOT
.pp
Invoke PUTBOOT with a command of the form:
.sp
.in 8
PUTBOOT [-H] <filename> <drive>
.fi
.sp
.in 0
where
.sp
.in 5
.ti -2
o -H is an optional flag discussed below;
.sp
.ti -2
o <filename> is the name of the file to be written to the system tracks;
.sp
.ti -2
o <drive> is the drive specifier for the drive to which <filename> is to be 
written (letter in the range A-P.)
.sp
.in 0
.pp
PUTBOOT writes the specified file to the system tracks of the specified 
drive.  Sector skewing is not used; the file is written to the system tracks 
in physical sector number order.  
.ix sector skewing
.pp
Because the file that is written is normally in command file format, 
PUTBOOT contains special logic to strip off the first 28 bytes of the file 
whenever the file begins with the number 601AH, the magic number used in 
command files.  If, by chance, the file to be written begins with 601AH, but 
should not have its first 28 bytes discarded, the -H flag should be specified 
in the PUTBOOT command line.  This flag tells PUTBOOT to write the file 
verbatim to the system tracks.
.ix -H flag
.pp
PUTBOOT uses BDOS calls to read <filename>, and used BIOS calls to
write <filename> to the system tracks.  It refers to the OFF and SPT 
parameters in the Disk Parameter Block to determine how large the system
track space is.  The source and command files for PUTBOOT are supplied
on the distribution disks for CP/M-8000.
.ix OFF parameter
.ix SPT parameter
.sp 2
.ce
End of Section 8
.bp
.he CP/M-8000 System Guide                            End of Section 8
.sp 23
.bp
.nx appa

