
MP/M-86 is created in four different stages:

1) Kernal and A86 files:

    A)	The A86 assembly modules that make up the operating system
	are assembled on an 8080, Z-80, or 8086 based micro using ASM86.
	GENCMD is run on each file and then they are renamed to type
	MPM.  These modules are:

	SUP.MPM, RTM.MPM, MEM.MPM, CIO.MPM, BDOS.MPM, XIOS.MPM

    B)	RSP files are like the MPM files above, they are renamed
	CMD files.  The H86 files are the output of ASM86.

	TMP.RSP, ECHO.RSP, CLOCK.RSP, SUBMIT.CMD, LDBDOS.H86, LDMPM.H86,
	LDBIOS.H86

    C)	Files on the distribution diskettes created by combining
	the above files.  The end user can also create these files.

	MPM.SYS, MPMLDR.CMD 
 

2) Utilities compiled on the VAX

	There are several directories on the VAX of utility source.
	Submit jobs, using files of type COM, are run on the vax
	to compile, link, located and 'H86' the utilities.
	These submit jobs can be initiated from a micro via
	submit jobs run under MP/M II - see the files
	with names such as COMALL.SUB or COMMIX.SUB.
	Once the utilities are compiled on the VAX,
	submits are run under MP/M II to upload and gencmd
	the H86 files - see the files with names such as GENALL.SUB,
	GENMIX.SUB.

	The utilities created in this manner are:

	ABORT.CMD, ATTACH.CMD, ASM86.CMD, CONSOLE.CMD, DIR.CMD
	DSKRESET.CMD, ED.CMD, ERA.CMD, ERAQ.CMD, GENSYS.CMD,
	MPMSTAT.CMD, MPMSTAT.RSP, PIP.CMD, REN.CMD, SDIR.CMD,
	SET.CMD, SHOW.CMD, SPOOL.CMD, STAT.CMD, STOPSPLR.CMD,
	TOD.CMD, TYPE.CMD
	

3) Utilities compiled on a micro

	These PL/M programs are compiled under CP/M running the ISIS
	interface.  They are:

	DDT86.CMD, GENCMD.CMD, GENDEF.CMD 


4) Text and source files included on the distribution diskettes:

	DEBLOCK.LIB, ROM.A86, SINGLES.DEF, SINGLES.LIB, SYSDAT.LIB,
	SYSTEM.LIB, TMP.A86, XIOS.A86, ROM.A86



Thå entirå sourcå foò MP/M-8¶ 2.° ió backeä uð oî tapå createä oî
thå VAX®  Thió includeó thå fileó thaô arå compileä oò assembleä
oî micros®  Thå entirå sourcå ió alsï duplicateä oî singlå
densitù ¸ incè floppies¬ writteî bù ALTOÓ machines®  Archievå
copieó oæ thå disketteó anä tapeó arå kepô iî thå Digitaì
Researcè banë vault.

Source and compiled and assembled list files are on the VAX in
sub-directories of DRB0:[FRANK.MPM86].
