.bp odd
.cs 5
.mt 5
.mb 6
.pl 66
.ll 65
.po 10
.pp 5
.hm 2
.fm 2
.he 
.ft All Information Presented Here is Proprietary to Digital Research
.ce 2
.tc F  New Functions and Implementation Changes
.sh
Appendix F
.sp
.sh
    New Functions and Implementation Changes
.sp 2
.pp 5
CP/M-8000 has six new Basic Disk Operating System (BDOS) functions
and additional implementation changes in the BDOS functions and
data structures that differ from other CP/M systems. 
.sp 2
.ce
.sh
Table F-1.  New BDOS Functions
.nf
.sp
.in 11
.ll 60
          Function                 Number
.sp
     Get Free Disk Space             46
.sp
     Chain to Program                47
.sp
     Flush Buffers                   48
.sp 
     Set Exception Vector            61
.sp
     Set Supervisor State            62
.sp
     Get/Set TPA Limits              63
.sp 2
.in 0
.fi
.tc    F.1  BDOS Function and Data Structure Changes
.he CP/M-8000 Programmer's Guide            F.1  BDOS Function Changes
.sh
F.1  BDOS Function and Data Structure Changes
.pp
Implementation changes in CP/M-8000 BDOS functions and data structures
are described in the following table:
.sp 2 
.sh
.ce
Table F-2.  BDOS Function Implementation Changes
.sp
.ll 60
.in 5
.nf
   BDOS Function       Number       Implementation  
                                        Change
.sp
.in 36
.ti -32
.fi
Return Version Number     12    Contains the version number 2022H indicating 
CP/M-8000 Version 1.1.
.sp
.ti -32
Reset Disk System         14    Does not log in disk drive A when it resets 
the disk system.
.sp
.ti -32
Open File                 15    Opens a file only at extent 0, the base extent.
.sp
.ti -32
Get Disk Parameters       31    Returns a copy of the Disk Parameter Block 
(DPB).
.in 0
.fi
.sp 2
.in 5
.nf
.sh
Table F-3.  BDOS Data Structure Implementation Changes
.qs
.sp
.ll 60
     Structure                  Implementation  
                                    Change
.fi
.sp
.in 28
.ti -23
Base Page              Additional information has been added.  The base page 
is no longer located at a fixed address.  Appendix C outlines the structure 
of the base page.
.sp
.ti -23
File Control Block     The byte sequence for the Random Record Field has 
changed.  The most significant byte (r0) is first and the least significant 
byte (r2) is last. 
.in 0
.sp 2
.tc    F.2  BDOS Functions Not Supported By CP/M-8000
.he CP/M-8000 Programmer's Guide       F.2  Unsupported BDOS Functions
.sh
F.2  BDOS Functions Not Supported By CP/M-8000 
.qs
.pp
The list below contains functions and commands supported by other CP/M
systems, but that are not supported by CP/M-8000. 
.sp 2
.ce
.sh
Table F-4.  BDOS Functions Not Supported by CP/M-8000
.qs
.nf
.sp
.ll 60
.in 7
         BDOS Function                   Number
.nf
.sp
.in 7 
Get Address of Allocation Vector           27
.sp
Set DMA Base+                              51
.sp
Get DMA Base+                              52
.sp
Get Maximum Memory*                        53
.sp
Get Absolute Memory*                       54
.sp
Allocate Absolute Memory*                  55
.sp
Free Memory*                               56
.sp
Free All Memory*                           57
.fi
.in 7
.sp 2
.ti -2
+ The Z8000 microprocessor does not have a segmented 
architecture.  Therefore, functions involving segment registers are not 
relevant to CP/M-8000.
.sp
.ti -2
* CP/M-8000 does not have memory management functions.
.in 0
.sp
.mb 5
.fm 1
.fi
.pp
DDT-8000 does not support the Assemble (A) command.
.sp 2
.ce
End of Appendix F
.nx index

